Perturbing SU($n$)-symmetric interactions

### 2020-01-17 collective_sun/sun_perturbation.tex
fixed typo in Eq. (25) added a proof (new appendix) of a non-vanishing spectral gap for power-law interactions with \alpha <= D

\newcommand{\EE}{\mathbb{E}}
\newcommand{\RR}{\mathbb{R}}
\renewcommand{\SS}{\mathbb{S}}
\newcommand{\VV}{\mathbb{V}}
\newcommand{\ZZ}{\mathbb{Z}}

@@

If the coefficients $h_{pq}$ of the interaction Hamiltonian $H_0$ are
all negative, then the ground-state manifold $\M_0$ of the interaction
Hamiltonian $H_0$ consists of fully symmetric states that are
simultaneous $+1$ eigenstates of all permutation operators $\Pi_{pq}$.
We will adopt the restriction that all $h_{pq}<0$ throughout these
notes, but note that this restriction can be relaxed to the assumption
that the initial state of the spins is fully symmetric (e.g.~a
spin-polarized state).  We will also assume throughout these notes
that the fully symmetric manifold $\M_0$ is gapped away from all other
states by an interaction energy difference that is large compared to
any coupling between $\M_0$ and its orthogonal complement.  The
precise conditions for such a non-vanishing many-body interaction
energy gap is an important question that deserves further study in
future work.

@@

If we {\it construct} a perturbation $\V_2^\Delta$ with coefficients
$\v{\m w}_O^\Delta$ that satisfy \eqref{eq:cond_2} with eigenvalue
$\Delta$, then
\begin{align}
  H_0 \V_2^\Delta \ket\psi = \p{E_0 + \Delta} \V_2^\Delta \ket\psi,
\end{align}
which implies that $\V_2^\Delta\ket\psi$ is an (unnormalized state of
definite energy that depends only on the coefficients $\m w_O^\Delta$.
Similarly to the case of single-body perturbations, we can decompose
any vector $\v{\m w}_O$ into its projections $\v{\m w}_O^\Delta$ onto
the eigenspace of vectors satisfying \eqref{eq:cond_2} with eigenvalue
$\Delta$, i.e.~$\v{\m w}_O=\sum_\Delta\v{\m w}_O^\Delta$, which allows
us to expand
\begin{align}
  \V_2 = \sum_\Delta \V_2^\Delta,
  &&
  \V_2^\Delta \equiv \sum_O \v{\m w}_O^\Delta \c \v{\m O},
\end{align}
and in turn write
\begin{align}
  H_2^{(2)} = - \P_0 \V_2^\Delta \E \V_2^\Delta \P_0
  = -\sum_{\Delta\ne0} \f1\Delta \P_0 \p{\V_2^\Delta}^2 \P_0,
\end{align}
where the product $\P_0 \p{\V_2^\Delta}^2 \P_0$ is simplified in
Appendices \ref{sec:POQP} and \ref{sec:multi_to_collective}.  In the
special case that
\begin{align}
  \V_2 = \sum_{p<q} w_{pq} s_\z^{(p)} s_\z^{(q)},
\end{align}
where $s_\z\equiv\p{\op\up-\op\dn}/2$ is an SU(2) spin-$z$ operator,
then
\begin{align}
  H_2^{(2)} \EQFS W_4 S_\z^4 - W_2 S_\z^2,
\end{align}
where $S_\z\equiv\sum_p s_\z^{(p)}$ is a collective SU(2) spin
operator, $W_k$ are scalar coefficients, and we have neglected scalar
terms in $H_2^{(2)}$ without physical consequence.  In order to expand
the coefficients $W_k$, for any ${N \choose 2}$-component vector
$\v{\m m}=\sum_{p<q}m_{pq}\ket{\set{p,q}}$ we define the $N$-component
vector
\begin{align}
  \v m \equiv \sum_{p,q} m_{pq} \ket{p},
\end{align}
where we define $m_{pq}$ for all $p,q$ (as opposed to only $p<q$) by
enforcing $m_{pq}=m_{qp}$ and $m_{pp}=0$.  We then denote the
projection of $\v{\m m}$ onto the $\Delta$-eigenspace of the matrix
$\check{\m h}$ defined in \eqref{eq:h_super_mat} by $\v{\m m}_\Delta$,
which allows us to expand
\begin{align}
  W_4
  \equiv \sum_{\Delta\ne0} \f1\Delta \f{\v w_\Delta\c\v w_\Delta
    - \v{\m w}_\Delta\c\v{\m w}_\Delta}{N!_4},
\end{align}
and
\begin{align}
  W_2
  \equiv \sum_{\Delta\ne0} \f1{\Delta} \sp{\f{\v w_\Delta\c\v w_\Delta
      - 2\v{\m w}_\Delta\c\v{\m w}_\Delta}{4N!_2}
    + \f{\p{3N-4}\p{\v w_\Delta\c\v w_\Delta
        - \v{\m w}_\Delta\c\v{\m w}_\Delta}}{2N!_4}},
\end{align}
where
\begin{align}
  N!_k \equiv \prod_{j=0}^{k-1} \p{N-j} = \f{N!}{\p{N-k}!}.
\end{align}

@@

@@

For reference, the single-body eigenvalue problem is
\begin{align}
  \p{\m h - \diag\v h}\c\v v = \Delta \v v,
  \label{eq:cond_1_ref}
\end{align}
where $\m h$ is a matrix of all couplings $h_{pq}$ that appear in the
interaction Hamiltonian $H_0$ in \eqref{eq:ints}; $\v h$ is a vector
of all $h_p\equiv\sum_q h_{pq}$; $\diag\v h$ is a matrix with $\v h$
on the diagonal and zeroes everywhere else; and $\v v$ is a vector of
coefficients used to construct the single-body perturbation
$\V_1\p{\v v}$ defined in \eqref{eq:perturbations}.  When the
interaction Hamiltonian $H_0$ is translationally invariant, the
couplings $h_{pq}$ depend only on the separation $\abs{p-q}$, so the
eigenvectors of $\m h$ are plane waves of the form
\begin{align}
  \v v_k \equiv \sum_p e^{ip\c k} \ket{p},
\end{align}
where on a $D$-dimensional periodic lattice of $N=L^D$ spins, lattice
sites are indexed by vectors $p\in\ZZ_L^D$, and wavenumbers take on
values $k\in\ZZ_L^D\times2\pi/N$.  The corresponding eigenvalues of
$\m h$ can be determined by expanding
\begin{align}
  \m h\c\v v_k = \sum_{p,q} h_{pq} e^{iq\c k} \ket{p}
  = \sum_{p,d} h_{p,d+p} e^{i\p{d+p}\c k} \ket{p}
  = \sum_d h_{0,d} e^{id\c k} \v v_k
  = \sum_d h_{0,d} \cos\p{d\c k} \v v_k,
\end{align}
where the imaginary terms vanish due to the fact that
$h_{0,d}=h_{0,-d}$.  The remainder of \eqref{eq:cond_1_ref} that we
need to sort out is $\diag\v h$, where all
$h_p=\sum_qh_{pq}=\sum_dh_{0,d}$ are equal, which implies that
$\diag\v h=\sum_dh_{0,d}$ is a scalar.  We thus find that
\begin{align}
  \p{\m h - \diag\v h}\c\v v_k = \Delta_k \v v_k,
  &&
  \Delta_k \equiv \sum_d h_{0,d} \sp{\cos\p{d\c k}-1}.
\end{align}

@@

The two-body eigenvalue equation in \eqref{eq:cond_2} is designed to
find solutions $\m w$ to the equation
\begin{align}
  \p{2h_{pq}-h_p-h_q} w_{pq}
  + \sum_{k\notin\set{p,q}} \p{h_{pk} w_{kq} + w_{pk} h_{kq}}
  = \Delta w_{pq},
  \label{eq:cond_2_ref_start}
\end{align}
where $h_p\equiv\sum_q h_{pq}$.  If the interaction Hamiltonian $H_0$
and perturbation $\V_2\p{\m w}$ are translationally invariant, then
the couplings $h_{pq}$ and coefficients $w_{pq}$ depend on $p$ and $q$
only through the separation $\abs{p-q}$.  In this case, all $h_p=h_0$
are equal, and \eqref{eq:cond_2_ref_start} can be written in the form
\begin{align}
  2 \m h\circ \m w + \m h\c\m w + \m w\c\m h
  = \sp{\Delta + 2h_0} \m w,
  \label{eq:cond_2_ref}
\end{align}
where $\m h$ and $\m w$ are matrices of all $h_{pq}$ and $w_{pq}$, and
$\m h\circ\m w$ denotes an element-wise (Kronecker) product of $\m h$
and $\m w$, i.e.~with matrix elements $h_{pq}w_{pq}$.  In order to
find solutions to \eqref{eq:cond_2_ref}, we we expand the matrices
$\m h,\m w$ in a basis that respects translational symmetry.  To this
end, for all lattice site displacements $d\in\ZZ_L^D$ we define the
{\it shift matrices}
\begin{align}
  \m s_d \equiv \sum_p \op{p+d}{p},
\end{align}
where the element-wise sums in $p+d$ are implicitly taken modulo $L$
(i.e.~the length of the periodic lattice along each axis).  The shift
matrices allow us to expand
\begin{align}
  \m h = \sum_{p,q} h_{pq} \op{p}{q}
  = \sum_{d,q} h_{q+d,q} \op{q+d}{q}
  = \sum_d h_{d,0} \m s_d,
  &&
  \m w = \sum_d w_{d,0} \m s_d,
\end{align}
and write the condition in \eqref{eq:cond_2_ref} as
\begin{align}
  \sum_{d\ne 0} h_{d,0} w_{d,0} \m s_d
  + \sum_{\substack{c,d\ne0\\c+d\ne0}} h_{c,0} w_{d,0} \m s_{c+d}
  = \p{\f12\Delta + h_0} \sum_{d\ne 0} w_{d,0} \m s_d,
\end{align}
where we explicitly exclude terms that always vanish due to the fact
that $\m h,\m w$ are zero on the diagonal.  With some re-indexing, we
can write
\begin{align}
  \sum_{d\ne 0} h_{d,0} w_{d,0} \m s_d
  + \sum_{\substack{c,d\ne0}} h_{d-c,0} w_{c,0} \m s_d
  = \p{\f12\Delta + h_0} \sum_{d\ne 0} w_{d,0} \m s_d,
  \label{eq:cond_2_shift}
\end{align}
and define
\begin{align}
  \ut{\check{\m h}}
  \equiv \sum_{d\ne0} h_{d,0} \op{d}
  + \sum_{c,d\ne0} h_{d-c,0} \op{d}{c},
  &&
  \ut{\v{\m w}} \equiv \sum_{d\ne0} w_{d,0} \ket{d},
\end{align}
in order to express \eqref{eq:cond_2_shift} as the eigenvalue
equation
\begin{align}
  \ut{\check{\m h}} \c \ut{\v{\m w}}
  = \p{\f12\Delta + h_0} \ut{\v{\m w}}.
  \label{eq:cond_2_square}
\end{align}
We can reduce this eigenvalue problem even further by making use of
reflection symmetries such as $h_{d,0}=h_{-d,0}$.  In $D$ dimensions,
there is one such symmetry for each lattice axis, which implies that
the degrees of freedom in $\ut{\v{\m w}}$ gets reduced by a factor of
$\sim 2^D$.  Reflection symmetries similarly reduce the degrees of
freedom in $\ut{\check{\m h}}$ from $\sim N\times N$ to
$\sim\p{N/2^D}\times\p{N/2^D}$.  We work out the case of $D=1$ below
as an example, and leave the general case of arbitrary dimension
$D\ge1$ to future work.

>>>>

\renewcommand{\SS}{\mathbb{S}}
\newcommand{\TT}{\mathbb{T}}
\newcommand{\VV}{\mathbb{V}}
\newcommand{\ZZ}{\mathbb{Z}}

@@

If the coefficients $h_{pq}$ of the interaction Hamiltonian $H_0$ are
all negative, then the ground-state manifold $\M_0$ of the interaction
Hamiltonian $H_0$ consists of fully symmetric states that are
simultaneous $+1$ eigenstates of all permutation operators $\Pi_{pq}$.
We will adopt the restriction that all $h_{pq}<0$ throughout these
notes, but note that this restriction can be relaxed to the assumption
that the initial state of the spins is fully symmetric (e.g.~a
spin-polarized state).  We will also assume throughout these notes
that the fully symmetric manifold $\M_0$ is gapped away from all other
states by an interaction energy difference that is large compared to
any coupling between $\M_0$ and its orthogonal complement.  Power-law
couplings of the form $h_{pq}=-h/\abs{p-q}^\alpha$ on a
$D$-dimensional lattice, for example, yield a non-vanishing spectral
gap in the thermodynamic limit when $\alpha\le D$ (see Appendix
\ref{sec:gap}).

@@

If we {\it construct} a perturbation $\V_2^\Delta$ with coefficients
$\v{\m w}_O^\Delta$ that satisfy \eqref{eq:cond_2} with eigenvalue
$\Delta$, then
\begin{align}
  H_0 \V_2^\Delta \ket\psi = \p{E_0 + \Delta} \V_2^\Delta \ket\psi,
\end{align}
which implies that $\V_2^\Delta\ket\psi$ is an (unnormalized state of
definite energy that depends only on the coefficients $\m w_O^\Delta$.
Similarly to the case of single-body perturbations, we can decompose
any vector $\v{\m w}_O$ into its projections $\v{\m w}_O^\Delta$ onto
the eigenspace of vectors satisfying \eqref{eq:cond_2} with eigenvalue
$\Delta$, i.e.~$\v{\m w}_O=\sum_\Delta\v{\m w}_O^\Delta$, which allows
us to expand
\begin{align}
  \V_2 = \sum_\Delta \V_2^\Delta,
  &&
  \V_2^\Delta \equiv \sum_O \v{\m w}_O^\Delta \c \v{\m O},
\end{align}
and in turn write
\begin{align}
  H_2^{(2)} = - \P_0 \V_2 \E \V_2 \P_0
  = -\sum_{\Delta\ne0} \f1\Delta \P_0 \p{\V_2^\Delta}^2 \P_0,
\end{align}
where the product $\P_0 \p{\V_2^\Delta}^2 \P_0$ is simplified in
Appendices \ref{sec:POQP} and \ref{sec:multi_to_collective}.  In the
special case that
\begin{align}
  \V_2 = \sum_{p<q} w_{pq} s_\z^{(p)} s_\z^{(q)},
\end{align}
where $s_\z\equiv\p{\op\up-\op\dn}/2$ is an SU(2) spin-$z$ operator,
the second-order effective Hamiltonian $H_2^{(2)}$ takes the form
\begin{align}
  H_2^{(2)} \EQFS W_4 S_\z^4 - W_2 S_\z^2,
\end{align}
where $S_\z\equiv\sum_p s_\z^{(p)}$ is a collective SU(2) spin
operator, $W_k$ are scalar coefficients, and we have neglected scalar
terms in $H_2^{(2)}$ without physical consequence.  In order to expand
the coefficients $W_k$, for any ${N \choose 2}$-component vector
$\v{\m m}=\sum_{p<q}m_{pq}\ket{\set{p,q}}$ we define the $N$-component
vector
\begin{align}
  \v m \equiv \sum_{p,q} m_{pq} \ket{p},
\end{align}
where we define $m_{pq}$ for all $p,q$ (as opposed to only $p<q$) by
enforcing $m_{pq}=m_{qp}$ and $m_{pp}=0$.  We then denote the
projection of $\v{\m m}$ onto the $\Delta$-eigenspace of the matrix
$\check{\m h}$ defined in \eqref{eq:h_super_mat} by $\v{\m m}_\Delta$,
which allows us to concisely write
\begin{align}
  W_4
  \equiv \sum_{\Delta\ne0} \f1\Delta \f{\v w_\Delta\c\v w_\Delta
    - \v{\m w}_\Delta\c\v{\m w}_\Delta}{N!_4},
\end{align}
and
\begin{align}
  W_2
  \equiv \sum_{\Delta\ne0} \f1{\Delta} \sp{\f{\v w_\Delta\c\v w_\Delta
      - 2\v{\m w}_\Delta\c\v{\m w}_\Delta}{4N!_2}
    + \f{\p{3N-4}\p{\v w_\Delta\c\v w_\Delta
        - \v{\m w}_\Delta\c\v{\m w}_\Delta}}{2N!_4}},
\end{align}
where
\begin{align}
  N!_k \equiv \prod_{j=0}^{k-1} \p{N-j} = \f{N!}{\p{N-k}!}
\end{align}
denotes a falling factorial.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Existence of a spectral gap with power-law interactions}
\label{sec:gap}

Here we show that power-law SU($n$)-symmetric interactions of the form
in \eqref{eq:ints} with $h_{pq}=-h/\abs{p-q}^\alpha$ yield a
non-vanishing many-body interaction energy gap when $\alpha<D$, where
$D$ is the dimension of the lattice.  On a periodic lattice of $N=L^D$
spins, the spectral gap of the interaction Hamiltonian is (see
Appendix \ref{sec:trans_inv_single})
\begin{align}
  \Delta_D
  = \sum_{d\in\tilde\ZZ_L^D} h_{0,d} \sp{\cos\p{d \c k_{\t{SE}}}-1}
  = h \sum_{\substack{d\in\tilde\ZZ_L^D\\d\ne0}}
  \f{1-\cos\p{d \c k_{\t{SE}}}}{\abs{d}^\alpha},
\end{align}
where the domain of $d$ is given by
$\tilde\ZZ_L\equiv\set{-\floor{L/2},-\floor{L/2}+1,\cdots,\ceil{L/2}}$,
and $k_{\t{SE}}$ is a wavenumber for a singly-excited spin-wave state.
In order to yield the smallest excitation energy $\Delta_D$, the
wavenumber $k_{\t{SE}}$ should maximize the contribution of the cosine
term above, which is achieved by a wavenumber that minimizes the
oscillations of this term when integrated over the entire lattice.  A
suitable candidate for a minimal wavenumber is
$k_{\t{SE}}=\p{2\pi/L,0,0,\cdots}$, which corresponds to an excitation
energy
\begin{align}
  \Delta_D = h \sum_{\substack{d\in\tilde\ZZ_L^D\\d\ne0}}
  \f{1-\cos\p{d_12\pi/L}}{\abs{d}^\alpha}.
\end{align}
Defining $\epsilon\equiv2/L$ and a rescaled domain
$\SS_\epsilon\equiv\tilde\ZZ_L/\epsilon\subset\sp{-1,1}$, we
substitute $x\equiv\epsilon d\in\SS_\epsilon^D$ to get
\begin{align}
  \Delta_D
  = h \sum_{\substack{x\in\SS_\epsilon^D\\x\ne0}}
  \f{1-\cos\p{\pi x_1}}{\abs{x/\epsilon}^\alpha}
  = h \epsilon^{D-\alpha} \sum_{\substack{x\in\SS_L^D\\x\ne0}}
  \epsilon^D \f{1-\cos\p{\pi x_1}}{\abs{x}^\alpha}.
\end{align}
As $\epsilon\to0$ the discrete sum over $x$ becomes an integral that
avoids an infinitesimal region about the origin, i.e.
\begin{align}
  \Delta_D = h \epsilon^{\alpha-D} \I_{D\epsilon},
  &&
  \I_{D\epsilon}
  \equiv \int_{\TT_1^D\setminus\TT_\epsilon^D} \d^Dx\,
  \f{1-\cos\p{\pi x_1}}{\abs{x}^\alpha},
\end{align}
where the integral $\I_{D\epsilon}$ is defined using the interval
$\TT_a\equiv\p{-a,a}$.  The integrand of $\I_{D\epsilon}$ is strictly
positive and well-behaved on the entirety of its domain except for the
origin, where depending on the value of $\alpha$ the integrand may
vanish or diverge as $\abs{x}\to0$.  Together, these facts mean that
\begin{align}
  \I_{D\epsilon} \stackrel{\epsilon\to0}{\sim} \epsilon^{-\gamma},
  &&
  \Delta_D \stackrel{\epsilon\to0}{\sim} h \epsilon^{\alpha-D-\gamma},
\end{align}
for some $\gamma\ge0$, which implies that gap $\Delta_D$ is
non-vanishing when
\begin{align}
  \alpha \le D + \gamma,
\end{align}
and in particular when $\alpha \le D$.

@@

For reference, the single-body eigenvalue problem is
\begin{align}
  \p{\m h - \diag\v h}\c\v v = \Delta \v v,
  \label{eq:cond_1_ref}
\end{align}
where $\m h$ is a matrix of all couplings $h_{pq}$ that appear in the
interaction Hamiltonian $H_0$ in \eqref{eq:ints}; $\v h$ is a vector
of all $h_p\equiv\sum_q h_{pq}$; $\diag\v h$ is a matrix with $\v h$
on the diagonal and zeroes everywhere else; and $\v v$ is a vector of
coefficients used to construct the single-body perturbation
$\V_1\p{\v v}$ defined in \eqref{eq:perturbations}.  When the
interaction Hamiltonian $H_0$ is translationally invariant, the
couplings $h_{pq}$ depend only on the separation $\abs{p-q}$, so the
eigenvectors of $\m h$ are plane waves of the form
\begin{align}
  \v v_k \equiv \sum_p e^{ip\c k} \ket{p},
\end{align}
where on a $D$-dimensional periodic lattice of $N=L^D$ spins, lattice
sites are indexed by vectors
$p\in\tilde\ZZ_L^D \equiv
\set{-\floor{L/2},-\floor{L/2}+1,\cdots,\ceil{L/2}}$, and wavenumbers
take on values $k\in\tilde\ZZ_L^D\times2\pi/L$.  The corresponding
eigenvalues of $\m h$ can be determined by expanding
\begin{align}
  \m h\c\v v_k = \sum_{p,q} h_{pq} e^{iq\c k} \ket{p}
  = \sum_{p,d} h_{p,d+p} e^{i\p{d+p}\c k} \ket{p}
  = \sum_d h_{0,d} e^{id\c k} \v v_k
  = \sum_d h_{0,d} \cos\p{d\c k} \v v_k,
\end{align}
where the imaginary terms vanish due to the fact that
$h_{0,d}=h_{0,-d}$.  The remainder of \eqref{eq:cond_1_ref} that we
need to sort out is $\diag\v h$, where all
$h_p=\sum_qh_{pq}=\sum_dh_{0,d}$ are equal, which implies that
$\diag\v h=\sum_dh_{0,d}$ is a scalar.  We thus find that
\begin{align}
  \p{\m h - \diag\v h}\c\v v_k = \Delta_k \v v_k,
  &&
  \Delta_k \equiv \sum_d h_{0,d} \sp{\cos\p{d\c k}-1}.
\end{align}

@@

The two-body eigenvalue equation in \eqref{eq:cond_2} is designed to
find solutions $\m w$ to the equation
\begin{align}
  \p{2h_{pq}-h_p-h_q} w_{pq}
  + \sum_{k\notin\set{p,q}} \p{h_{pk} w_{kq} + w_{pk} h_{kq}}
  = \Delta w_{pq},
  \label{eq:cond_2_ref_start}
\end{align}
where $h_p\equiv\sum_q h_{pq}$.  If the interaction Hamiltonian $H_0$
and perturbation $\V_2\p{\m w}$ are translationally invariant, then
the couplings $h_{pq}$ and coefficients $w_{pq}$ depend on $p$ and $q$
only through the separation $\abs{p-q}$.  In this case, all $h_p=h_0$
are equal, and \eqref{eq:cond_2_ref_start} can be written in the form
\begin{align}
  2 \m h\circ \m w + \m h\c\m w + \m w\c\m h
  = \sp{\Delta + 2h_0} \m w,
  \label{eq:cond_2_ref}
\end{align}
where $\m h$ and $\m w$ are matrices of all $h_{pq}$ and $w_{pq}$, and
$\m h\circ\m w$ denotes an element-wise (Kronecker) product of $\m h$
and $\m w$, i.e.~with matrix elements $h_{pq}w_{pq}$.  In order to
find solutions to \eqref{eq:cond_2_ref}, we we expand the matrices
$\m h,\m w$ in a basis that respects translational symmetry.  To this
end, for all lattice site displacements $d\in\tilde\ZZ_L^D$ we define
the {\it shift matrices}
\begin{align}
  \m s_d \equiv \sum_p \op{p+d}{p},
\end{align}
where the element-wise sums in $p+d$ are implicitly taken modulo $L$
(i.e.~the length of the periodic lattice along each axis).  The shift
matrices allow us to expand
\begin{align}
  \m h = \sum_{p,q} h_{pq} \op{p}{q}
  = \sum_{d,q} h_{q+d,q} \op{q+d}{q}
  = \sum_d h_{d,0} \m s_d,
  &&
  \m w = \sum_d w_{d,0} \m s_d,
\end{align}
and write the condition in \eqref{eq:cond_2_ref} as
\begin{align}
  \sum_{d\ne 0} h_{d,0} w_{d,0} \m s_d
  + \sum_{\substack{c,d\ne0\\c+d\ne0}} h_{c,0} w_{d,0} \m s_{c+d}
  = \p{\f12\Delta + h_0} \sum_{d\ne 0} w_{d,0} \m s_d,
\end{align}
where we explicitly exclude terms that always vanish due to the fact
that $\m h,\m w$ are zero on the diagonal.  With some re-indexing, we
can write
\begin{align}
  \sum_{d\ne 0} h_{d,0} w_{d,0} \m s_d
  + \sum_{\substack{c,d\ne0}} h_{d-c,0} w_{c,0} \m s_d
  = \p{\f12\Delta + h_0} \sum_{d\ne 0} w_{d,0} \m s_d,
  \label{eq:cond_2_shift}
\end{align}
and define
\begin{align}
  \ut{\check{\m h}}
  \equiv \sum_{d\ne0} h_{d,0} \op{d}
  + \sum_{c,d\ne0} h_{d-c,0} \op{d}{c},
  &&
  \ut{\v{\m w}} \equiv \sum_{d\ne0} w_{d,0} \ket{d},
\end{align}
in order to express \eqref{eq:cond_2_shift} as the eigenvalue
equation
\begin{align}
  \ut{\check{\m h}} \c \ut{\v{\m w}}
  = \p{\f12\Delta + h_0} \ut{\v{\m w}}.
  \label{eq:cond_2_square}
\end{align}
We can reduce this eigenvalue problem even further by making use of
reflection symmetries such as $h_{d,0}=h_{-d,0}$.  In $D$ dimensions,
there is one such symmetry for each lattice axis, which implies that
the degrees of freedom in $\ut{\v{\m w}}$ gets reduced by a factor of
$\sim 2^D$.  Reflection symmetries similarly reduce the degrees of
freedom in $\ut{\check{\m h}}$ from $\sim N\times N$ to
$\sim\p{N/2^D}\times\p{N/2^D}$.  We work out the case of $D=1$ below
as an example, and leave the general case of arbitrary dimension
$D\ge1$ to future work.
